### 2020-04-30 LaTeX/sections/abstract.tex
New code (static array) + test cases

% !TEX root = ../main.tex

\begin{abstract}

Abstract goes here

\end{abstract}