### 2020-09-14 LaTeX/tables/eval.tex

% !TEX root = ../main.tex

\begin{table}[t]
\centering
\begin{tabular}{|l|c|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}

\multicolumn{1}{c}{\textbf{Exchange Platform}} & 
\multicolumn{1}{c}{\textbf{Examples}} & 
\headrow{Less Censorable} & 
\headrow{Minimum/ No  Trust} & 
\headrow{Fast Updates (No Latency)} & 
\headrow{High Performance} &
\headrow{Low Cost} &
\headrow{All ERC20 Compatible} &
\headrow{No Custody of Order Book} &
\headrow{No Custody of Funds} &
\headrow{Accurate Price Discovery} &
\headrow{High Transparency} &
\headrow{Automatic Matching} \\

\hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = == = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
\multirow{2}{*}{DEX w/ On-chain Order Book }      	& Bitshares		&\multirow{2}{*}{\full }		&\multirow{2}{*}{\full }		&\multirow{2}{*}{\prt }		&\multirow{2}{*}{half }		&\multirow{2}{*}{half }		&\multirow{2}{*}{half } 		&\multirow{2}{*}{\full }		&\multirow{2}{*}{\full }		&\multirow{2}{*}{\full }		&\multirow{2}{*}{\full }		&\multirow{2}{*}{\full } 		\\ 													
											&Stellar			& 						&						&						&						&						&						&						&						&						& 						&						\\  \hline						
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = == = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
\multirow{3}{*}{DEX w/ Off-chain Order Book}        &0x				&\multirow{3}{*}{half}		&\multirow{3}{*}{half }		&\multirow{3}{*}{\full }		&\multirow{3}{*}{\full }		&\multirow{3}{*}{\full }		&\multirow{3}{*}{\full } 		&\multirow{3}{*}{\prt }		&\multirow{3}{*}{\full }		&\multirow{3}{*}{\full }		&\multirow{3}{*}{\prt }		&\multirow{3}{*}{half } 		\\ 
							                      	&EtherDelta		& 						&						&						&						&						&						&						&						&						& 						&						\\
										       &IDEX     		& 						&						&						&						&						&						&						&						&						& 						&						\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = == = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
\multirow{3}{*}{Automated Market Making DEX}    &UniSwap		&\multirow{3}{*}{\full}		&\multirow{3}{*}{half }		&\multirow{3}{*}{ DN}		&\multirow{3}{*}{half}		&\multirow{3}{*}{ half}		&\multirow{3}{*}{half} 		&\multirow{3}{*}{ \full}		&\multirow{3}{*}{ \prt}		&\multirow{3}{*}{\full }		&\multirow{3}{*}{ \full}		&\multirow{3}{*}{ DN} 		\\ 
							                      	&Bancor			& 						&						&						&						&						&						&						&						&						& 						&						\\
										       &Curve     		& 						&						&						&						&						&						&						&						&						& 						&						\\ \hline			
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = == = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
\multirow{3}{*}{CEX}    							&Binance		&\multirow{3}{*}{\prt}		&\multirow{3}{*}{\prt}		&\multirow{3}{*}{\full}		&\multirow{3}{*}{\full}		&\multirow{3}{*}{\full}		&\multirow{3}{*}{\full} 		&\multirow{3}{*}{\full}		&\multirow{3}{*}{\prt}		&\multirow{3}{*}{\prt}		&\multirow{3}{*}{\prt }		&\multirow{3}{*}{\full } 		\\ 
							                      	&Coinbase		& 						&						&						&						&						&						&						&						&						& 						&						\\
										       &Bitfinex     		& 						&						&						&						&						&						&						&						&						& 						&						\\ \hline	
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = == = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
\multicolumn{2}{|c|}{Auction Based Model (us)}       & 						&						&						&						&						&						&						&						&						& 						&						\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = == = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{Comparative evaluation of different types of exchange platforms: {\full} indicates the properties (columns) are fulfilled by the corresponding mechanism (rows) within reason, {\prt} means the property is not fulfilled, \textblue{say sth about half dot when you figure out how to generate it}}
\label{tab:eval}}
\end{table}