### 2020-09-14 LaTeX/sections/abstract.tex
more edits

% !TEX root = ../main.tex

\begin{abstract}
Order-driven exchanges on Ethereum tend to operate in a fully centralized fashion, or with some functionality performed on-chain (\eg loading accounts, order cancelation) and some off-chain functionality (\eg matching orders). Fully on-chain solutions have a superior threat model and were experimented with in the early days of Ethereum, but they have been largely abandoned for performance reasons. As blockchain performance continues to improve (both in research and in practice) and on-chain dealer-based exchanges like Uniswap (an alternative to order-based exchanges) rise in popularity, it seems to be an ideal time to revisit the idea of a fully on-chain order-driven exchange and benchmark what is possible performance-wise. We design and optimize an Ethereum-based call market exchange, \cm, and conduct a variety of experiments to show the current benchmark is around one hundred trade executions per block. We also explore numerous extensions for further improvements.
\end{abstract}