### 2021-01-11 LaTeX-USENIX/sections/appendix.tex
more random edits

% !TEX root = ../main.tex

%% = = = = = = = = = = = =Transferring the Collateral back to the traders = = = = = = = = = = = = =  %
%\subsection{Collateralization}
%
%Options for non-collateralized markets
%ERC20 to ERC20 -> extra transaction
%