### 2021-08-26 LaTeX-LNCS-2022/main.tex
cutting the text down to 23 pages

\documentclass[envcountsect]{llncs}
\pagestyle{headings} 
\bibliographystyle{abbrv}

% = = = = = Packages = = = = = %

\input{setup/packages}
\input{setup/macros}
\input{setup/table}

%Get out of jail free:
%\usepackage[all=normal,floats=tight,paragraphs=tight]{savetrees}

% = = = = = Title = = = = = %

\begin{document}

\frontmatter
\mainmatter

\title{\Large \bf TBD}

%\author{Mahsa Moosavi\inst{1} \and Jeremy Clark\inst{1}}
%\institute{Concordia University, Montreal, Canada}

\maketitle

% = = = = = Abstract = = = = = %

\input{sections/abstract}

% = = = = = Main Body = = = = = %

\input{sections/paper}

% = = = = = Bibliography = = = = = %

\clearpage
\bibliography{bib/DEX}
%\nocite{*}

% = = = = = End Notes = = = = = %

\clearpage
\appendix
\input{sections/appendix.tex}

\end{document}