### 2021-08-30 LaTeX-LNCS-2022/tables/tradeexecutions.tex
Entire "trade execution system" section rewritten

% !TEX root = ../main.tex

% = = = = = = = = = = = =Comparison Trade Executions = = =  = = = = = = = = = = = %
\begin{table}[t]
%\setlength{\tabcolsep}{1pt}% Shrink \tabcolsep
\centering
\scriptsize
%\begin{tabular}{|c|c|c|c|c|}
\begin{tabular}{|>{\centering}m{4.5cm} |>{\centering}m{1cm} |>{\centering}m{1cm} |>{\centering}m{1cm} |>{\centering\arraybackslash}m{1cm}|}

\multicolumn{1}{c}{\scriptsize{Trade Execution System}}  & \headrowcleaning{\scriptsize {High Performance}} & \headrowcleaning{\scriptsize {High Security}} & \headrowcleaning{\scriptsize {Easy to Regulate}} & \headrowcleaning{\scriptsize {Easy to Realize}}\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Centralized Exchange (CEX)}       		 & \footnotesize\Smiley     	&\footnotesize\Sadey 		&\footnotesize\Smiley 		& \footnotesize\Smiley                 \\ \hline
	\textbf{Decentralized Exchange (DEX)}		 &\footnotesize\Smiley 		&\footnotesize\Sadey 		&\footnotesize\Smiley 		&\footnotesize\Smiley		\\ \hline
	\textbf{Fully On-chain Exchange}			 &\footnotesize\Sadey 		&\footnotesize\Smiley 		&\footnotesize\Sadey 		&\footnotesize\Sadey					\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
\end{tabular}
\caption{\footnotesize{}.
\label{tab:exchanges}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = =  = = = = = = = = = = =  = = = = = = = = = == = = =  = =%