### 2021-09-10 LaTeX-LNCS-2022/sections/abstract.tex
top matter [19 pages total]

% !TEX root = ../main.tex

\begin{abstract}

Financial regulators have long-standing concerns about fully decentralized exchanges that run `on-chain' without any obvious regulatory hooks. Uniswap's open protocol and the next generation of such automated market makers (AMMs) made these concerns a reality. AMMs implement a lightweight dealer-based trading system, but it is unlike anything on Wall Street, requires fees intrinsically, and is susceptible to front-running attacks. This leaves the following research questions we address in this paper: (i) are conventional (\ie orderbooks) and secure (\ie resistant to front-running and price manipulation) exchanges feasible as a fully on-chain service on a public blockchain, (ii) what is the performance profile, and (iii) how much do Layer 2 techniques (\eg Arbitrum) increase performance? To answer these questions, we implement, benchmark, and experiment with an Ethereum-based call market (or batch auction) exchange called \cm, demonstrating many of Ethereum's nuances. We confirm the functionality is too heavy for Ethereum today (you cannot expect to exceed a few hundred trade executions per block) but show it scales dramatically (99.88\% gas cost reduction) on Arbitrum.

\end{abstract}