### 2021-09-13 LaTeX-LNCS-2022/tables/eval2.tex
hooks for improvements

% !TEX root = ../main.tex
% = = = = = = = = = = = = = = = = = = = = =Call market operations Table = = = = = = = = = = = =  = = = = = =  %
\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{|p{1.5cm}|p{3cm}|p{3.5cm}|p{3.5cm}|}

\hline
\textbf{Type}			 &\textbf{Description}   				& \textbf{Advantages}      				& \textbf{Disadvantages}                   \\ \hline

% = = = = = = = = = = = =%
Centralized Exchanges (CEX)
& Order-driven exchange acts as a trusted third party (\eg Binance, Bitfinex)
& Conventional \newline
Highest performance \newline
Low fees \newline
Easy to regulate \newline
Low price slippage  \newline
Verbose trading strategies 
&  Fully trusted custodian \newline
Slow withdrawals \newline
Server downtime \newline
Uncertain fair execution 
\\ 
\hline

% = = = = = = = = = = = =%
Partially On-chain Exchange
& Order-driven exchange acts as a semi-trusted (\eg EtherDelta, 0x, IDEX, Loopring)
&  High performance \newline
Low fees \newline
Easy to regulate \newline
Low price slippage \newline
Verbose trading strategies \newline
Semi-custodial
& 
Slow withdrawals \newline
Server downtime \newline
Front-running attacks \newline
Uncertain fair execution
\\ 
\hline

% = = = = = = = = = = = =%
On-Chain Dealers
& Quote-driven decentralized exchange trades from inventory with public pricing rule (\eg Uniswap v3)
&  Non-custodial \newline
Instant trading \newline
Moderate performance \newline
Fair execution 
&  Unconventional \newline
Impermanent loss \newline
High price slippage \newline
Intrinsic fees \newline
Front-running attacks \newline
Limited trading strategies \newline
Hard to regulate
\\ 
\hline

% = = = = = = = = = = = =%
On-chain Order-Driven Exchanges 
& Order-driven decentralized exchange executes trades between buyers and sellers (\eg \cm)
&  Conventional \newline
Non-custodial \newline
Low price slippage \newline
Fair execution \newline
Verbose trading strategies \newline
Front-running is mitigable 
& Very low performance \newline
Hard to regulate
\\ 
\hline

\end{tabular}
\caption{ Comparison among different trade execution systems.
\label{tab:eval2}}
\end{table}

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %