### 2022-06-01 LaTeX-LNCS-2022/main.tex
wtsc changes

\documentclass[envcountsect]{llncs}
\pagestyle{headings} 
\bibliographystyle{abbrv}

% = = = = = Packages = = = = = %

\input{setup/packages}
\input{setup/macros}
\input{setup/table}

%Get out of jail free:
%\usepackage[all=normal,floats=tight,paragraphs=tight]{savetrees}

% = = = = = Title = = = = = %

\begin{document}

\frontmatter
\mainmatter

\title{\Large \bf Lissy: Experimenting with on-chain order books}

\author{Mahsa Moosavi \and Jeremy Clark}
\institute{Concordia University, Montreal, Canada \\ \email{seye\_moo@encs.concordia.ca}}

\maketitle

% = = = = = Abstract = = = = = %

\input{sections/abstract}

% = = = = = Main Body = = = = = %

\input{sections/paper}

% = = = = = Acknowledgements = = = = = %

\subsubsection*{Acknowledgements.}  The authors thank the AMF (Autorité des Marchés Financiers) for supporting this research project. J. Clark also acknowledges partial funding from the National Sciences and Engineering Research Council (NSERC)/Raymond Chabot Grant Thornton/Catallaxy Industrial Research Chair in Blockchain Technologies, as well as NSERC through a Discovery Grant. M. Moosavi acknowledges support from Fonds de Recherche du Québec - Nature et Technologies (FRQNT).

% = = = = = Bibliography = = = = = %

\bibliography{bib/DEX}
%\nocite{*}

% = = = = = End Notes = = = = = %

\appendix
\input{sections/appendix.tex}

\end{document}